---
title: "Market Segmentation for Airlines"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(caret)
```

# About problem  

Market segmentation is a strategy that divides a broad target market of customers into smaller, more similar groups, and then designs a marketing strategy specifically for each group. Clustering is a common technique for market segmentation since it automatically finds similar groups given a data set. 
  
In this problem, we'll see how clustering can be used to find similar groups of customers who belong to an airline's frequent flyer program. The airline is trying to learn more about its customers so that it can target different customer segments with different types of mileage offers.  
  
# About data  
  
The file AirlinesCluster.csv contains information on 3,999 members of the frequent flyer program. This data comes from the textbook "Data Mining for Business Intelligence," by Galit Shmueli, Nitin R. Patel, and Peter C. Bruce. For more information, see the website for the book.  
  
There are seven different variables in the dataset, described below:  
  
*Balance* = number of miles eligible for award travel  

*QualMiles* = number of miles qualifying for TopFlight status  

*BonusMiles* = number of miles earned from non-flight bonus transactions in the past 12 months  

*BonusTrans* = number of non-flight bonus transactions in the past 12 months  

*FlightMiles* = number of flight miles in the past 12 months  

*FlightTrans* = number of flight transactions in the past 12 months  

*DaysSinceEnroll* = number of days since enrolled in the frequent flyer program  

# Problem 1.1 - Normalizing the Data

Read the dataset AirlinesCluster.csv into R and call it "airlines".
```{r}
airlines <- read.csv("AirlinesCluster.csv")
```

*Q1: Looking at the summary of airlines, which TWO variables have (on average) the smallest values?*
```{r}
summary(airlines)
```
*A:* FlightTrans, BonusTrans.  
  
*Q2: Which TWO variables have (on average) the largest values?*  
*A:* Balance, BonusMiles.  
  
# Problem 1.2 - Normalizing the Data  
  
In this problem, we will normalize our data before we run the clustering algorithms.  
*Q3: Why is it important to normalize the data before clustering?*  
*A:* If we don't normalize the data, the clustering will be dominated by the variables that are on a larger scale.  

# Problem 1.3 - Normalizing the Data

Let's go ahead and normalize our data. You can normalize the variables in a data frame by using the preProcess function in the "caret" package.  Now, create a normalized data frame called "airlinesNorm" by running the following commands:
```{r}
preproc <- preProcess(airlines)
airlinesNorm <- predict(preproc, airlines)
```

The first command pre-processes the data, and the second command performs the normalization. If you look at the summary of airlinesNorm, you should see that all of the variables now have mean zero. You can also see that each of the variables has standard deviation 1 by using the sd() function.
```{r}
summary(airlinesNorm)
```

*Q4: In the normalized data, which variable has the largest maximum value?*  
*A:* FlightMiles.  
  
*Q5: In the normalized data, which variable has the smallest minimum value?*  
*A:* DaysSinceEnroll.  

# Problem 2.1 - Hierarchical Clustering

Compute the distances between data points (using euclidean distance) and then run the Hierarchical clustering algorithm (using method="ward.D") on the normalized data. It may take a few minutes for the commands to finish since the dataset has a large number of observations for hierarchical clustering.
```{r}
distances <- dist(airlinesNorm, method = "euclidean")
```
Hierarchical clustering
```{r}
clusterAirlines <- hclust(distances, method = "ward.D") 
```
Then, plot the dendrogram of the hierarchical clustering process.  
```{r}
plot(clusterAirlines)
```

Suppose the airline is looking for somewhere between 2 and 10 clusters.  
*Q6: According to the dendrogram, which of the following is NOT a good choice for the number of clusters?*  
*A:* 6.  
*Explanation.* If you run a horizontal line down the dendrogram, you can see that there is a long time that the line crosses 2 clusters, 3 clusters, or 7 clusters. However, it it hard to see the horizontal line cross 6 clusters. This means that 6 clusters is probably not a good choice.  
# Problem 2.2 - Hierarchical Clustering  

Suppose that after looking at the dendrogram and discussing with the marketing department, the airline decides to proceed with 5 clusters. Divide the data points into 5 clusters by using the cutree function.  
```{r}
clusterGroups <- cutree(clusterAirlines, k = 5)
```

*Q7: How many data points are in Cluster 1?*  
```{r}
table(clusterGroups)
```

# Problem 2.3 - Hierarchical Clustering

Now, use tapply to compare the average values in each of the variables for the 5 clusters (the centroids of the clusters). You may want to compute the average values of the unnormalized data so that it is easier to interpret. You can do this for the variable "Balance" with the following command:
```{r}
tapply(airlines$Balance, clusterGroups, mean)
```
QualMiles
```{r}
tapply(airlines$QualMiles, clusterGroups, mean)
```

```{r}
tapply(airlines$BonusMiles, clusterGroups, mean)
```

```{r}
tapply(airlines$BonusTrans, clusterGroups, mean)
```

```{r}
tapply(airlines$FlightMiles, clusterGroups, mean)
```

```{r}
tapply(airlines$FlightTrans, clusterGroups, mean)
```

```{r}
tapply(airlines$DaysSinceEnroll, clusterGroups, mean)
```

*Q8: Compared to the other clusters, Cluster 1 has the largest average values in which variables (if any)? Select all that apply.*  
*A:* DaysSinceEnroll.  
  
*Q9: How would you describe the customers in Cluster 1?*  
*A:* Number of days since enrolled in the frequent flyer program  
  
*Explanation.* Cluster 1 mostly contains customers with few miles, but who have been with the airline the longest.  
  
# Problem 2.4 - Hierarchical Clustering

*Q10: Compared to the other clusters, Cluster 2 has the largest average values in which variables (if any)? Select all that apply.*  
*A: FlightTrans, FlightMiles, QualMiles.  
  
*Q11: How would you describe the customers in Cluster 2?*  
*A:* Customers who have accumulated a large amount of miles, and the ones with the largest number of flight transactions.  

# Problem 2.5 - Hierarchical Clustering  

*Q12: Compared to the other clusters, Cluster 3 has the largest average values in which variables (if any)? Select all that apply.*
*A:* BonusTrans, BonusMiles, Balance.  
  
*Q13: How would you describe the customers in Cluster 3?*  
*A:* Customers who have accumulated a large amount of miles, mostly through non-flight transactions.  

# Problem 2.6 - Hierarchical Clustering

*Q14: Compared to the other clusters, Cluster 4 has the largest average values in which variables (if any)? Select all that apply.*  
*A:* None.  
  
*Q15: How would you describe the customers in Cluster 4?*  
*A:* Relatively new customers who seem to be accumulating miles, mostly through non-flight transactions.  
*Explanation.* Cluster 4 customers have the smallest value in DaysSinceEnroll, but they are already accumulating a reasonable number of miles.  

# Problem 2.7 - Hierarchical Clustering

*Q16: Compared to the other clusters, Cluster 5 has the largest average values in which variables (if any)? Select all that apply.*  
*A:* Cluster 5 does not have the largest values in any of the variables.  
  
*Q17: How would you describe the customers in Cluster 5?*  
*A:* Relatively new customers who don't use the airline very often.  

# Problem 3.1 - K-Means Clustering

Now run the k-means clustering algorithm on the normalized data, again creating 5 clusters. Set the seed to 88 right before running the clustering algorithm, and set the argument iter.max to 1000.
```{r}
set.seed(88)
KMC <- kmeans(airlinesNorm, centers = 5, iter.max = 1000)
str(KMC)
```

*Q18: How many clusters have more than 1,000 observations?*  
```{r}
KMC$size
```

# Problem 3.2 - K-Means Clustering  

Now, compare the cluster centroids to each other either by dividing the data points into groups and then using tapply, or by looking at the output of kmeansClust\$centers, where "kmeansClust" is the name of the output of the kmeans function. (Note that the output of kmeansClust$centers will be for the normalized data. If you want to look at the average values for the unnormalized data, you need to use tapply like we did for hierarchical clustering.)
```{r}
KMC$centers
```

*Q19: Do you expect Cluster 1 of the K-Means clustering output to necessarily be similar to Cluster 1 of the Hierarchical clustering output?*

1) Yes, because the clusters are displayed in order of size, so the largest cluster will always be first.  

2) Yes, because the clusters are displayed according to the properties of the centroid, so the cluster order will be similar.  

3) No, because cluster ordering is not meaningful in either k-means clustering or hierarchical clustering (correct answer).  
*Explanation* The clusters are not displayed in a meaningful order, so while there may be a cluster produced by the k-means algorithm that is similar to Cluster 1 produced by the Hierarchical method, it will not necessarily be shown first.  

4) No, because the clusters produced by the k-means algorithm will never be similar to the clusters produced by the Hierarchical algorithm.  

