---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

In this problem, we will be examining the "state" dataset, which has data from the 1970s on all fifty US states. For each state, the dataset includes the population, per capita income, illiteracy rate, murder rate, high school graduation rate, average number of frost days, area, latitude and longitude, division the state belongs to,  region the state belongs to, and two-letter abbreviation.

Load data:
```{r}
state_data <- read.csv("data/statedata.csv")
```
Let's view our data:
```{r}
glimpse(state_data)
```

This dataset has 50 observations (one for each US state) and the following 15 variables:

Population - the population estimate of the state in 1975
Income - per capita income in 1974
Illiteracy - illiteracy rates in 1970, as a percent of the population
Life.Exp - the life expectancy in years of residents of the state in 1970
Murder - the murder and non-negligent manslaughter rate per 100,000 population in 1976 
HS.Grad - percent of high-school graduates in 1970
Frost - the mean number of days with minimum temperature below freezing from 1931–1960 in the capital or a large city of the state
Area - the land area (in square miles) of the state
state.abb - a 2-letter abreviation for each state
state.area - the area of each state, in square miles
x - the longitude of the center of the state
y - the latitude of the center of the state
state.division - the division each state belongs to (New England, Middle Atlantic, South Atlantic, East South Central, West South Central, East North Central, West North Central, Mountain, or Pacific)
state.name - the full names of each state
state.region - the region each state belong to (Northeast, South, North Central, or West)

# Problem 1.1 - Data Exploration

We begin by exploring the data. Plot all of the states' centers with latitude on the y axis (the "y" variable in our dataset) and longitude on the x axis (the "x" variable in our dataset). The shape of the plot should look like the outline of the United States! Note that Alaska and Hawaii have had their coordinates adjusted to appear just off of the west coast.

```{r}
ggplot(state_data, aes(x, y)) +
  geom_point()
```

## Problem 1.2 - Data Exploration

Determine which region of the US (West, North Central, South, or Northeast) has the highest average high school graduation rate of all the states in the region:
```{r}
state_data %>% 
  group_by(state.region) %>% 
  summarise(Avg = mean(HS.Grad)) %>% 
  arrange(desc(Avg))
```
## Problem 1.3 - Data Exploration

Now, make a boxplot of the murder rate by region (for more information about creating boxplots in R, type ?boxplot in your console).

Which region has the highest median murder rate?
```{r}
ggplot(state_data, aes(x = state.region, y = Murder)) + 
  geom_boxplot()
```
## Problem 1.4 - Data Exploration 
You should see that there is an outlier in the Northeast region of the boxplot you just generated. Which state does this correspond to? (Hint: There are many ways to find the answer to this question, but one way is to use the subset command to only look at the Northeast data.)
```{r}
state_data %>% 
  dplyr::filter(state.region == "Northeast") %>% 
  arrange(desc(Murder)) %>% 
  select(state.name, Murder)
```

## Problem 2.1 - Predicting Life Expectancy - An Initial Model

We would like to build a model to predict life expectancy by state using the state statistics we have in our dataset.

Build the model with all potential variables included (Population, Income, Illiteracy, Murder, HS.Grad, Frost, and Area). Note that you should use the variable "Area" in your model, NOT the variable "state.area".

What is the coefficient for "Income" in your linear regression model?
```{r}
LifeExp_model <- lm(Life.Exp ~ Population + Income + Illiteracy + Murder + HS.Grad + Frost + Area, data = state_data)
summary(LifeExp_model)
```
## Problem 2.2 - Predicting Life Expectancy - An Initial Model

Call the coefficient for income x (the answer to Problem 2.1). What is the interpretation of the coefficient x?
```{r}
LifeExp_model$coefficients
```
**Answer:** for a one unit increase in income, predicted life expectancy decreases by |x|
**Explanation:** If we increase income by one unit, then our model’s prediction will increase by the coefficient of income, x. Because x is negative, this is the same as predicted life expectancy decreasing by |x|.

## Problem 2.3 - Predicting Life Expectancy - An Initial Model

Now plot a graph of life expectancy vs. income.
Visually observe the plot. What appears to be the relationship?
```{r}
ggplot(state_data, aes(Income, Life.Exp)) + 
  geom_point()
```
**Answer:** life expectancy is somewhat positive correlated wiht income.

## Problem 2.4 - Predicting Life Expectancy - An Initial Model

The model we built does not display the relationship we saw from the plot of life expectancy vs. income. Which of the following explanations seems the most reasonable?

**Answer:** multicollinearity

**Explanation:** Although income is an insignificant variable in the model, this does not mean that there is no association between income and life expectancy. However, in the presence of all of the other variables, income does not add statistically significant explanatory power to the model. This means that multicollinearity is probably the issue.

## Problem 3.1 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Recall that we discussed the principle of simplicity: that is, a model with fewer variables is preferable to a model with many unnnecessary variables. Experiment with removing independent variables from the original model. Remember to use the significance of the coefficients to decide which variables to remove (**remove the one with the largest "p-value" first, or the one with the "t value" closest to zero**), and to remove them one at a time (this is called **"backwards variable selection"**). This is important due to multicollinearity issues - removing one insignificant variable may make another previously insignificant variable become significant.

You should be able to find a good model with only 4 independent variables, instead of the original 7. Which variables does this model contain?

Let's overview base model LifeExp_model once again:
```{r}
summary(LifeExp_model)
```
First get rid of Area variable
```{r}
LifeExp_model2 <- lm(Life.Exp ~ Population + Income + Illiteracy + Murder + HS.Grad + Frost, data = state_data)
summary(LifeExp_model2)
```
Second, remove Illiteracy:
```{r}
LifeExp_model3 <- lm(Life.Exp ~ Population + Income + Murder + HS.Grad + Frost, data = state_data)
summary(LifeExp_model3)
```
Third, remove Income
```{r}
LifeExp_model4 <- lm(Life.Exp ~ Population + Murder + HS.Grad + Frost, data = state_data)
summary(LifeExp_model4)
```
**Explanation:** We would eliminate the variable "Area" first (since it has the highest p-value, or probability, with a value of 0.9649), by adjusting our lm command to the following:

LinReg = lm(Life.Exp ~ Population + Income + Illiteracy + Murder + HS.Grad + Frost, data=statedata)

Looking at summary(LinReg) now, we would choose to eliminate "Illiteracy" since it now has the highest p-value of 0.9340, using the following command:

LinReg = lm(Life.Exp ~ Population + Income + Murder + HS.Grad + Frost, data=statedata)

Looking at summary(LinReg) again, we would next choose to eliminate "Income", since it has a p-value of 0.9153. This gives the following four variable model:

LinReg = lm(Life.Exp ~ Population + Murder + HS.Grad + Frost, data=statedata)

This model with 4 variables is a good model. However, we can see that the variable "Population" is not quite significant. In practice, it would be up to you whether or not to keep the variable "Population" or eliminate it for a 3-variable model. Population does not add much statistical significance in the presence of murder, high school graduation rate, and frost days. However, for the remainder of this question, we will analyze the 4-variable model.

Fourt, remove Population
```{r}
LifeExp_model5 <- lm(Life.Exp ~ Murder + HS.Grad + Frost, data = state_data)
summary(LifeExp_model5)
```

## Problem 3.2 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Removing insignificant variables changes the Multiple R-squared value of the model. By looking at the summary output for both the initial model (all independent variables) and the simplified model (only 4 independent variables) and using what you learned in class, which of the following correctly explains the change in the Multiple R-squared value?

**Answer:** We expect the "Multiple R-squared" value of the simplified model to be slightly worse than that of the initial model. It can't be better than the "Multiple R-squared" value of the initial model.

**Explanation:**When we remove insignificant variables, the "Multiple R-squared" will always be worse, but only slightly worse. This is due to the nature of a linear regression model. It is always possible for the regression model to make a coefficient zero, which would be the same as removing the variable from the model. The fact that the coefficient is not zero in the intial model means it must be helping the R-squared value, even if it is only a very small improvement. So when we force the variable to be removed, it will decrease the R-squared a little bit. However, this small decrease is worth it to have a simpler model.

On the contrary, when we remove insignificant variables, the "Adjusted R-squred" will frequently be better. This value accounts for the complexity of the model, and thus tends to increase as insignificant variables are removed, and decrease as insignificant variables are added.

## Problem 3.3 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Using the simplified 4 variable model that we created, we'll now take a look at how our predictions compare to the actual values.

Take a look at the vector of predictions by using the predict function (since we are just looking at predictions on the training set, you don't need to pass a "newdata" argument to the predict function).

**Question:**Which state do we predict to have the lowest life expectancy? (Hint: use the sort function)

```{r}
LifExPredictions <- predict(LifeExp_model4)
```
Answer:
```{r}
which.min(LifExPredictions)[[1]] %>% 
  state_data$state.name[.]
```

**Explanation:**If your simplified 4-variable model is called "LinReg", you can answer this question by typing sort(predict(LinReg)) in your R console. The first state listed has the lowest predicted life expectancy, which is Alabama.

**Question:**Which state actually has the lowest life expectancy? (Hint: use the which.min function)

```{r}
#state_data[which.min(state_data$Life.Exp), "state.name"]
which.min(state_data$Life.Exp) %>% 
  state_data[., "state.name"]
```

## Problem 3.4 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

**Question:**Which state do we predict to have the highest life expectancy?

```{r}
which.max(LifExPredictions)[[1]] %>% 
  state_data$state.name[.]
```

**Question:**Which state actually has the highest life expectancy?

```{r}
which.max(state_data$Life.Exp) %>% 
  state_data[., "state.name"]
```

## Problem 3.5 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Take a look at the vector of residuals (the difference between the predicted and actual values).

**Question:**For which state do we make the smallest absolute error?

```{r}
which.min(abs(state_data$Life.Exp - LifExPredictions))[[1]] %>% 
  state_data$state.name[.]
```

**Question**For which state do we make the largest absolute error?

```{r}
which.max(abs(state_data$Life.Exp - LifExPredictions))[[1]] %>% 
  state_data$state.name[.]
```
