---
title: "Separating Spam from Ham (Part 1)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tm)
library(caTools)
library(rpart)
library(rpart.plot)
library(randomForest)
library(ROCR)
library(magrittr)
```

# About problem  
  
Nearly every email user has at some point encountered a "spam" email, which is an unsolicited message often advertising a product, containing links to malware, or attempting to scam the recipient. Roughly 80-90% of more than 100 billion emails sent each day are spam emails, most being sent from botnets of malware-infected computers. The remainder of emails are called "ham" emails.  
  
As a result of the huge number of spam emails being sent across the Internet each day, most email providers offer a spam filter that automatically flags likely spam messages and separates them from the ham. Though these filters use a number of techniques (e.g. looking up the sender in a so-called "Blackhole List" that contains IP addresses of likely spammers), most rely heavily on the analysis of the contents of an email via text analytics.  
  
In this homework problem, we will build and evaluate a spam filter using a publicly available dataset first described in the 2006 conference paper "Spam Filtering with Naive Bayes -- Which Naive Bayes?" by V. Metsis, I. Androutsopoulos, and G. Paliouras. The "ham" messages in this dataset come from the inbox of former Enron Managing Director for Research Vincent Kaminski, one of the inboxes in the Enron Corpus. One source of spam messages in this dataset is the SpamAssassin corpus, which contains hand-labeled spam messages contributed by Internet users. The remaining spam was collected by Project Honey Pot, a project that collects spam messages and identifies spammers by publishing email address that humans would know not to contact but that bots might target with spam. The full dataset we will use was constructed as roughly a 75/25 mix of the ham and spam messages.  
  
# About data

The dataset contains just two fields:  

*text:* The text of the email.  

*spam:* A binary variable indicating if the email was spam.  
  
# Problem 1.1 - Loading the Dataset  

Begin by loading the dataset emails.csv into a data frame called emails. Remember to pass the stringsAsFactors=FALSE option when loading the data.  
```{r}
emails <- read.csv("emails.csv", stringsAsFactors = F)
str(emails)
```

*Q1: How many emails are in the dataset?*  
```{r}
nrow(emails)
```

# Problem 1.2 - Loading the Dataset  

*Q2: How many of the emails are spam?*  
```{r}
table(emails$spam)
```

# Problem 1.3 - Loading the Dataset  
  
*Q3: Which word appears at the beginning of every email in the dataset?*  
Respond as a lower-case word with punctuation removed.  
*A:* subject.  

# Problem 1.4 - Loading the Dataset  

*Q4: Could a spam classifier potentially benefit from including the frequency of the word that appears in every email?*  
*A:* Yes -- the number of times the word appears might help us differentiate spam from ham.  

# Problem 1.5 - Loading the Dataset  
  
The nchar() function counts the number of characters in a piece of text.  
*Q5: How many characters are in the longest email in the dataset (where longest is measured in terms of the maximum number of characters)?*  
```{r}
max(nchar(emails$text))
```

# Problem 1.6 - Loading the Dataset  
  
*Q6: Which row contains the shortest email in the dataset?*  
(Just like in the previous problem, shortest is measured in terms of the fewest number of characters.)
```{r}
which.min(nchar(emails$text))
```

# Problem 2.1 - Preparing the Corpus  

Follow the standard steps to build and pre-process the corpus:  

1) Build a new corpus variable called corpus:
```{r}
corpus <- VCorpus(VectorSource(emails$text))
corpus[[1]]$content
```

2) Using tm_map, convert the text to lowercase:
```{r}
corpus <- tm_map(corpus, content_transformer(tolower))
corpus[[1]]$content
```

3) Using tm_map, remove all punctuation from the corpus:
```{r}
corpus <- tm_map(corpus, removePunctuation)
corpus[[1]]$content
```

4) Using tm_map, remove all English stopwords from the corpus:
```{r}
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus[[1]]$content
```

5) Using tm_map, stem the words in the corpus:
```{r}
corpus <- tm_map(corpus, stemDocument)
corpus[[1]]$content
```

6) Build a document term matrix from the corpus, called dtm:
```{r}
dtm <- DocumentTermMatrix(corpus)
dtm
```

*Q7: How many terms are in dtm?*  
*A:* 28687.  
  
# Problem 2.2 - Preparing the Corpus  

To obtain a more reasonable number of terms, limit dtm to contain terms appearing in at least 5% of documents, and store this result as spdtm (don't overwrite dtm, because we will use it in a later step of this homework).  
```{r}
spdtm <- removeSparseTerms(dtm, 0.95)
spdtm
```
*Q8: How many terms are in spdtm?*  
*A:* 330.  

# Problem 2.3 - Preparing the Corpus  

Build a data frame called emailsSparse from spdtm, and use the make.names function to make the variable names of emailsSparse valid.  
```{r}
(emailsSparse <- as.data.frame(as.matrix(spdtm)))
```
```{r}
colnames(emailsSparse) <- make.names(colnames(emailsSparse))
emailsSparse
```

colSums() is an R function that returns the sum of values for each variable in our data frame. Our data frame contains the number of times each word stem (columns) appeared in each email (rows). Therefore, colSums(emailsSparse) returns the number of times a word stem appeared across all the emails in the dataset.  
*Q9: What is the word stem that shows up most frequently across all the emails in the dataset?*  
```{r}
freq <- colSums(emailsSparse)
which.max(freq)
```
Hint: think about how you can use sort() or which.max() to pick out the maximum frequency.  
  
# Problem 2.4 - Preparing the Corpus  

Add a variable called "spam" to emailsSparse containing the email spam labels. You can do this by copying over the "spam" variable from the original data frame (remember how we did this in the Twitter lecture).  
```{r}
emailsSparse$spam <- emails$spam
```

*Q:10 How many word stems appear at least 5000 times in the ham emails in the dataset?*
```{r}
(stem5000 <- freq[freq >= 5000])
names(stem5000)
```
```{r}
# Filter dataframe from not spam and replace 'span' column
ham_df <- subset(emailsSparse, spam == 0)[, !(names(emailsSparse) == 'spam')]
freq2 <- colSums(ham_df)
freq2[freq2 >= 5000]
```
*A:* 5.  

# Problem 2.5 - Preparing the Corpus  

*Q11: How many word stems appear at least 1000 times in the spam emails in the dataset?*  
```{r}
spam_df <- subset(emailsSparse, spam == 1)[, !(names(emailsSparse) == 'spam')]
freq3 <- colSums(spam_df)
freq3[freq3 >= 1000]
```

# Problem 2.6 - Preparing the Corpus  

*Q12: The lists of most common words are significantly different between the spam and ham emails. What does this likely imply?*  

*A:* The frequencies of these most common words are likely to help differentiate between spam and ham.  

# Problem 2.7 - Preparing the Corpus  

Several of the most common word stems from the ham documents, such as "enron", "hou" (short for Houston), "vinc" (the word stem of "Vince") and "kaminski", are likely specific to Vincent Kaminski's inbox.  
*Q13 :What does this mean about the applicability of the text analytics models we will train for the spam filtering problem?*  

*A:* The models we build are personalized, and would need to be further tested before being used as a spam filter for another person.  
  
# Problem 3.1 - Building machine learning models  

First, convert the dependent variable to a factor:
```{r}
emailsSparse$spam = as.factor(emailsSparse$spam)
```

Next, set the random seed to 123 and use the sample.split function to split emailsSparse 70/30 into a training set called "train" and a testing set called "test". Make sure to perform this step on emailsSparse instead of emails.  
```{r}
set.seed(123)
spl <- sample.split(emailsSparse$spam, 0.7)
train <- subset(emailsSparse, spl == TRUE)
test <- subset(emailsSparse, spl == FALSE)
```

Using the training set, train the following three machine learning models. The models should predict the dependent variable "spam", using all other available variables as independent variables. Please be patient, as these models may take a few minutes to train.  
  
1) A logistic regression model called spamLog. You may see a warning message here - we'll discuss this more later.  
```{r}
spamLog <- glm(spam ~ ., data = train, family = "binomial")
```

2) A CART model called spamCART, using the default parameters to train the model (don't worry about adding minbucket or cp). Remember to add the argument method="class" since this is a binary classification problem.  
```{r}
spamCART <- rpart(spam ~ ., data = train, method = "class")
```

3) A random forest model called spamRF, using the default parameters to train the model (don't worry about specifying ntree or nodesize). Directly before training the random forest model, set the random seed to 123 (even though we've already done this earlier in the problem, it's important to set the seed right before training the model so we all obtain the same results. Keep in mind though that on certain operating systems, your results might still be slightly different).  
```{r}
set.seed(123)
spamRF <- randomForest(spam ~ ., data = train, method = "class")
```

For each model, obtain the predicted spam probabilities for the training set. Be careful to obtain probabilities instead of predicted classes, because we will be using these values to compute training set AUC values. Recall that you can obtain probabilities for CART models by not passing any type parameter to the predict() function, and you can obtain probabilities from a random forest by adding the argument type="prob". For CART and random forest, you need to select the second column of the output of the predict() function, corresponding to the probability of a message being spam.  

1) Predict probabilities logistic regression:
```{r}
predTrainLog <- predict(spamLog, newdata = train, type = "response")
#predLog[1:10]
#probLog <- exp(predLog)/(1 + exp(predLog))
#probLog[1:10]
```

2) Predict CART probabilities:
```{r}
predTrainCART <- predict(spamCART, newdata = train)
predTrainCART[1:10,]
```

3) Predict randome forest probabilities:
```{r}
predTrainRF <- predict(spamRF, newdata = train, type="prob")
predTrainRF[1:10,]
```

You may have noticed that training the logistic regression model yielded the messages "algorithm did not converge" and "fitted probabilities numerically 0 or 1 occurred". Both of these messages often indicate overfitting and the first indicates particularly severe overfitting, often to the point that the training set observations are fit perfectly by the model. Let's investigate the predicted probabilities from the logistic regression model.

*Q14: How many of the training set predicted probabilities from spamLog are less than 0.00001?*
```{r}
table(predTrainLog < 0.00001)
```

*Q15: How many of the training set predicted probabilities from spamLog are more than 0.99999?*
```{r}
table(predTrainLog > 0.99999)
```

*Q16: How many of the training set predicted probabilities from spamLog are between 0.00001 and 0.99999?*  
```{r}
table(predTrainLog >= 0.00001 & predTrainLog <= 0.99999)
```

# Problem 3.2 - Building Machine Learning Models  

*Q17: How many variables are labeled as significant (at the p=0.05 level) in the logistic regression summary output?*  
*A:* 0.  

# Problem 3.3 - Building Machine Learning Models  

*Q18: How many of the word stems "enron", "hou", "vinc", and "kaminski" appear in the CART tree?*  
Recall that we suspect these word stems are specific to Vincent Kaminski and might affect the generalizability of a spam filter built with his ham data.
```{r}
prp(spamCART)
```

*A:* 2.  

# Problem 3.4 - Building Machine Learning Models  

*Q19: What is the training set accuracy of spamLog, using a threshold of 0.5 for predictions?*
```{r}
table(train$spam, predTrainLog > 0.5)
```
Accuracy:
```{r}
(3052 + 954) / nrow(train)
```

# Problem 3.5 - Building Machine Learning Models  

*Q20: What is the training set AUC of spamLog?*
```{r}
predROCRlog <- prediction(predTrainLog, train$spam)
```

Compute AUC
```{r}
as.numeric(performance(predROCRlog, "auc")@y.values)
```

# Problem 3.6 - Building Machine Learning Models  

*Q21: What is the training set accuracy of spamCART, using a threshold of 0.5 for predictions?*  
(Remember that if you used the type="class" argument when making predictions, you automatically used a threshold of 0.5. If you did not add in the type argument to the predict function, the probabilities are in the second column of the predict output.)
```{r}
table(train$spam, predTrainCART[,2] > 0.5)
```
Accuracy:
```{r}
(2885 + 894) / nrow(train)
```

# Problem 3.7 - Building Machine Learning Models

*Q22: What is the training set AUC of spamCART?*  
(Remember that you have to pass the prediction function predicted probabilities, so don't include the type argument when making predictions for your CART model.)
```{r}
predROCRcart <- prediction(predTrainCART[,2], train$spam)
```
Compute AUC
```{r}
as.numeric(performance(predROCRcart, "auc")@y.values)
```

# Problem 3.8 - Building Machine Learning Models

*Q23: What is the training set accuracy of spamRF, using a threshold of 0.5 for predictions?*  
(Remember that your answer might not match ours exactly, due to random behavior in the random forest algorithm on different operating systems.)
```{r}
table(train$spam, predTrainRF[,2] > 0.5)
```
Accuracy:
```{r}
(3046 + 958) / nrow(train)
```

# Problem 3.9 - Building Machine Learning Models  

*Q24: What is the training set AUC of spamRF?*  
(Remember to pass the argument type="prob" to the predict function to get predicted probabilities for a random forest model. The probabilities will be the second column of the output.)
```{r}
predROCRrf <- prediction(predTrainRF[,2], train$spam)
```
Compute AUC
```{r}
as.numeric(performance(predROCRrf, "auc")@y.values)
```

# Problem 3.10 - Building Machine Learning Models

*Q25: Which model had the best training set performance, in terms of accuracy and AUC?*
*A:* logistic regression.  

# Problem 4.1 - Evaluating on the Test Set  

Obtain predicted probabilities for the testing set for each of the models, again ensuring that probabilities instead of classes are obtained.

*Q26: What is the testing set accuracy of spamLog, using a threshold of 0.5 for predictions?*
```{r}
predTestLog <- predict(spamLog, newdata = test, type = "response")
table(test$spam, predTestLog > 0.5)
```
Accuracy:
```{r}
(1257 + 376) / nrow(test)
```

# Problem 4.2 - Evaluating on the Test Set

*Q27: What is the testing set AUC of spamLog?*
```{r}
predROCRlogTest <- prediction(predTestLog, test$spam)
```

Compute AUC
```{r}
as.numeric(performance(predROCRlogTest, "auc")@y.values)
```

# Problem 4.3 - Evaluating on the Test Set  

*Q28: What is the testing set accuracy of spamCART, using a threshold of 0.5 for predictions?* 
```{r}
predTestCART <- predict(spamCART, newdata = test)
table(test$spam, predTestCART[,2] > 0.5)
```
Accuracy:
```{r}
(1228 + 386) / nrow(test)
```

# Problem 4.4 - Evaluating on the Test Set

*Q29: What is the testing set AUC of spamCART?*
```{r}
predROCRcartTest <- prediction(predTestCART[,2], test$spam)
```
Compute AUC
```{r}
as.numeric(performance(predROCRcartTest, "auc")@y.values)
```

# Problem 4.5 - Evaluating on the Test Set  

*Q30: What is the testing set accuracy of spamRF, using a threshold of 0.5 for predictions?*
```{r}
predTestRF <- predict(spamRF, newdata = test, type="prob")
table(test$spam, predTestRF[,2] > 0.5)
```
Accuracy:
```{r}
(1291 + 387) / (nrow(test))
```

# Problem 4.6 - Evaluating on the Test Set  

*Q31: What is the testing set AUC of spamRF?*
```{r}
predROCRrfTest <- prediction(predTestRF[,2], test$spam)
```
Compute AUC
```{r}
as.numeric(performance(predROCRrfTest, "auc")@y.values)
```

# Problem 4.7 - Evaluating on the Test Set

*Q32: Which model had the best testing set performance, in terms of accuracy and AUC?*
*A:* Random Forest.  
*Explanation.* The random forest outperformed logistic regression and CART in both measures, obtaining an impressive AUC of 0.997 on the test set.

# Problem 4.8 - Evaluating on the Test Set

*Q33: Which model demonstrated the greatest degree of overfitting?*  
*A:* Logistic regression.  
*Explanation.* Both CART and random forest had very similar accuracies on the training and testing sets. However, logistic regression obtained nearly perfect accuracy and AUC on the training set and had far-from-perfect performance on the testing set. This is an indicator of overfitting.

# Problem 5.1 - Assigning weights to different types of errors  

Thus far, we have used a threshold of 0.5 as the cutoff for predicting that an email message is spam, and we have used accuracy as one of our measures of model quality. As we have previously learned, these are good choices when we have no preference for different types of errors (false positives vs. false negatives), but other choices might be better if we assign a higher cost to one type of error.  
  
Consider the case of an email provider using the spam filter we have developed. The email provider moves all of the emails flagged as spam to a separate "Junk Email" folder, meaning those emails are not displayed in the main inbox. The emails not flagged as spam by the algorithm are displayed in the inbox. Many of this provider's email users never check the spam folder, so they will never see emails delivered there.  
  
*Q34: In this scenario, what is the cost associated with the model making a false negative error?*  

1) A ham email will be sent to the Junk Email folder, potentially resulting in the email user never seeing that message.  

2) A spam email will be displayed in the main inbox, a nuisance for the email user (correct)  

*Explanation.* A false negative means the model labels a spam email as ham. This results in a spam email being displayed in the main inbox.

3) There is no cost associated with this sort of mistake.

*Q35: In this scenario, what is the cost associated with our model making a false positive error?*  

1) A ham email will be sent to the Junk Email folder, potentially resulting in the email user never seeing that message (correct).  

*Explanation.* A false positive means the model labels a ham email as spam. This results in a ham email being sent to the Junk Email folder.  

2) A spam email will be displayed in the main inbox, a nuisance for the email user.  

3) There is no cost associated with this sort of mistake.  

# Problem 5.2 - Assigning Weights to Different Types of Errors  

*Q36: Which sort of mistake is more costly (less desirable), assuming that the user will never check the Junk Email folder?*  

1) False negative  

2) False positive (correct).
  
*Explanation*. A false negative is largely a nuisance (the user will need to delete the unsolicited email). However a false positive can be very costly, since the user might completely miss an important email due to it being delivered to the spam folder. Therefore, the false positive is more costly.  
  
# Problem 5.3 - Assigning Weights to Different Types of Errors  
  
*Q37: What sort of user might assign a particularly high cost to a false negative result?*  

1) A user who does not mind spam emails reaching their main inbox.  

2) A user who is particularly annoyed by spam email reaching their main inbox  (correct).  

*Explanation.*A false negative results in spam reaching a user's main inbox, which is a nuisance. A user who is particularly annoyed by such spam would assign a particularly high cost to a false negative.  

3) A user who never checks their Junk Email folder  

4) A user who always checks their Junk Email folder  
  
# Problem 5.4 - Assigning Weights to Different Types of Errors

*Q38: What sort of user might assign a particularly high cost to a false positive result?*  

3) A user who never checks their Junk Email folder  

*Explanation.* A false positive results in ham being sent to a user's Junk Email folder. While the user might catch the mistake upon checking the Junk Email folder, users who never check this folder will miss the email, incurring a particularly high cost.  

# Problem 5.5 - Assigning Weights to Different Types of Errors

Consider another use case for the spam filter, in which messages labeled as spam are still delivered to the main inbox but are flagged as "potential spam." Therefore, there is no risk of the email user missing an email regardless of whether it is flagged as spam. What is the largest way in which this change in spam filter design affects the costs of false negative and false positive results?
  
1) The cost of false negative results is decreased  

2) The cost of false negative results is increased  

3) The cost of false positive results is decreased (correct)  

*Explanation.* While before many users would completely miss a ham email labeled as spam (false positive), now users will not miss an email after this sort of mistake. As a result, the cost of a false positive has been decreased.  

4) The cost of false positive results is increased

# Problem 5.6 - Assigning Weights to Different Types of Errors  

Consider a large-scale email provider with more than 100,000 customers. Which of the following represents an approach for approximating each customer's preferences between a false positive and false negative that is both practical and personalized?  

1) Use the expert opinion of a project manager to select the relative cost for all users  

2) Automatically collect information about how often each user accesses his/her Junk Email folder to infer preferences (correct).  

*Explanation.* While using expert opinion is practical, it is not personalized (we would use the same cost for all users). Likewise, a random sample of user preferences doesn't enable personalized costs for each user.  
While a survey of all users would enable personalization, it is impractical to obtain survey results from all or most of the users.  
While it's impractical to survey all users, it is easy to automatically collect their usage patterns. This could enable us to select higher regression thresholds for users who rarely check their Junk Email folder but lower thresholds for users who regularly check the folder.  

3) Survey a random sample of users to measure their preferences  

4) Survey all users to measure their preferences  

# Problem 6.1 - Integrating Word Count Information

While we have thus far mostly dealt with frequencies of specific words in our analysis, we can extract other information from text. The last two sections of this problem will deal with two other types of information we can extract.  
  
First, we will use the number of words in the each email as an independent variable. We can use the original document term matrix called dtm for this task. The document term matrix has documents (in this case, emails) as its rows, terms (in this case word stems) as its columns, and frequencies as its values. As a result, the sum of all the elements in a row of the document term matrix is equal to the number of terms present in the document corresponding to the row. Obtain the word counts for each email with the command:
```{r}
wordCount <- rowSums(as.matrix(dtm))
```
When you have successfully created wordCount, answer the following question.  

*Q39: What would have occurred if we had instead created wordCount using spdtm instead of dtm?*  
```{r}
dtm
```
```{r}
spdtm
```

1) wordCount would have only counted some of the words and it would have only returned a result for some of the emails  

2) wordCount would have counted all of the words, but would have only returned a result for some the emails  

3) wordCount would have only counted some of the words, but would have returned a result for all the emails (correct)  

*Explanation.* spdtm has had sparse terms removed, which means we have removed some of the columns but none of the rows from dtm. This means rowSums will still return a sum for each row (one for each email), but it will not have counted the frequencies of any uncommon words in the dataset. As a result, wordCount will only count some of the words.  

4) wordCount would have counted all the words and it would have returned a result for all the emails  

# Problem 6.2 - Integrating Word Count Information  

Use the hist() function to plot the distribution of wordCount in the dataset.
```{r}
hist(wordCount)
```

*Q40: What best describes the distribution of the data?*

*A:* The data is skew right -- there are a large number of small wordCount values and a small number of large values.  

# Problem 6.3 - Integrating Word Count Information  

Now, use the hist() function to plot the distribution of log(wordCount) in the dataset.  
```{r}
hist(log(wordCount))
```

*Q41: What best describes the distribution of the data?*
*A:* The data is not skewed -- there are roughly the same number of unusually large and unusually small log(wordCount) values.  

# Problem 6.4 - Integrating Word Count Information  

Create a variable called logWordCount in emailsSparse that is equal to log(wordCount). 
```{r}
emailsSparse$logWordCount <- log(wordCount)
```

Use the boxplot() command to plot logWordCount against whether a message is spam.  
```{r}
boxplot(emailsSparse$logWordCount, emailsSparse$spam)
```

*Q42: Which of the following best describes the box plot?*  
*A:* logWordCount is slightly smaller in spam messages than in ham messages.  

*Explanation.* We can see that the 1st quartile, median, and 3rd quartiles are all slightly lower for spam messages than for ham messages.  

# Problem 6.5 - Integrating Word Count Information  

Because logWordCount differs between spam and ham messages, we hypothesize that it might be useful in predicting whether an email is spam. Take the following steps:  
  
1) Use the same sample.split output you obtained earlier (do not re-run sample.split) to split emailsSparse into a training and testing set, which you should call train2 and test2.
```{r}
train2 <- subset(emailsSparse, spl == TRUE)
test2 <- subset(emailsSparse, spl == FALSE)
```

2) Use train2 to train a CART tree with the default parameters, saving the model to the variable spam2CART.
```{r}
spam2CART <- rpart(spam ~ ., data = train2, method = "class")
```

3) Use train2 to train a random forest with the default parameters, saving the model to the variable spam2RF. Again, set the random seed to 123 directly before training spam2RF.
```{r}
set.seed(123)
spam2RF <- randomForest(spam ~ ., data = train2, method = "class")
```

*Q43: Was the new variable used in the new CART tree spam2CART?*  
*A:* Yes.  

# Problem 6.6 - Integrating Word Count Information  

Perform test-set predictions using the new CART and random forest models.  
CART model.
```{r}
predSpam2CART <- predict(spam2CART, newdata = test2)
table(test2$spam, predSpam2CART[,2] > 0.5)
```
Random Forest:
```{r}
predSpam2RF <- predict(spam2RF, newdata = test2, type="prob")
table(test2$spam, predSpam2RF[,2] > 0.5)
```

*Q44: What is the test-set accuracy of spam2CART, using threshold 0.5 for predicting an email is spam?*  
```{r}
(1214 + 384) / nrow(test2)
```

# Problem 6.7 - Integrating Word Count Information  

*Q45: What is the test-set AUC of spam2CART?*  
```{r}
ROCRspam2CART <- prediction(predSpam2CART[,2], test2$spam)
```
Compute AUC
```{r}
as.numeric(performance(ROCRspam2CART, "auc")@y.values)
```

# Problem 6.8 - Integrating Word Count Information  

*Q46: What is the test-set accuracy of spam2RF, using a threshold of 0.5 for predicting if an email is spam?*  
(Remember that you might get a different accuracy than us even if you set the seed, due to the random behavior of randomForest on some operating systems.)  
```{r}
(1298 + 381) / nrow(test2)
```

# Problem 6.9 - Integrating Word Count Information

*Q47: What is the test-set AUC of spam2RF?*  
(Remember that you might get a different AUC than us even if you set the seed when building your model, due to the random behavior of randomForest on some operating systems.)
```{r}
ROCRspam2RF <- prediction(predSpam2RF[,2], test2$spam)
```
Compute AUC
```{r}
as.numeric(performance(ROCRspam2RF, "auc")@y.values)
```
  
In this case, adding the logWordCounts variable did not result in improved results on the test set for the CART or random forest model.
