---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(magrittr)
library(purrr)
```

# In this problem, we'll use some basic data analysis in R to understand the motor vehicle thefts in Chicago. 

Here is a list of descriptions of the variables:

*ID*: a unique identifier for each observation  
*Date*: the date the crime occurred  
*LocationDescription*: the location where the crime occurred  
*Arrest*: whether or not an arrest was made for the crime (TRUE if an arrest was  
made, and FALSE if an arrest was not made)  
*Domestic*: whether or not the crime was a domestic crime, meaning that it was  
committed against a family member (TRUE if it was domestic, and FALSE if it was not domestic)  
*Beat*: the area, or "beat" in which the crime occurred. This is the smallest  
regional division defined by the Chicago police department.  
*District*: the police district in which the crime occured. Each district is  
composed of many beats, and are defined by the Chicago Police Department.  
*CommunityArea*: the community area in which the crime occurred.  
Since the 1920s, Chicago has been divided into what are called "community areas",  
of which there are now 77. The community areas were devised in an attempt to  
create socially homogeneous regions.  
*Year*: the year in which the crime occurred.  
*Latitude*: the latitude of the location at which the crime occurred.  
*Longitude*: the longitude of the location at which the crime occurred.  

# Loadint the data

```{r}
df <- read.csv("data/mvtWeek1.csv")
```
## Q1: How many rows of data (observations) are in this dataset?

```{r}
nrow(df)
```

## Q2: How many variables are in this dataset?

```{r}
ncol(df)
```

## Q3: Using the "max" function, what is the maximum value of the variable "ID"?

```{r}
max(df$ID)
```

## Q4: What is the minimum value of the variable "Beat"?

```{r}
min(df$Beat)
```

## Q5: How many observations have value TRUE in the Arrest variable  
## (this is the number of crimes for which an arrest was made)?

```{r}
table(df$Arrest)
```

## Q6: How many observations have a LocationDescription value of ALLEY?

```{r}
df %>%
  count(LocationDescription) %>% 
  filter(LocationDescription == 'ALLEY')
```

## 2.1. Understanding dates in R

Take a look at the first entry of Date (remember to use square brackets  
when looking at a certain entry of a variable).

## Q2.1: In what format are the entries in the variable Date?
```{r}
df$Date[100]
```
Now, let's convert these characters into a Date object in R.  
In your R console type:  
  
```{r}
DateConvert <- as.Date(strptime(df$Date, "%m/%d/%y %H:%M"))  
```

This converts the variable "Date" into a Date object in R.  
Take a look at the variable DateConvert using the summary function.  

## Q2.2: What is the month and year of the median date in our dataset?
## Enter your answer as "Month Year", without the quotes. (Ex: if the answer  
## was 2008-03-28, you would give the answer "March 2008", without the quotes.)

```{r}
summary(DateConvert)
```

Now, let's extract the month and the day of the week, and add these variables to  
our data frame mvt. We can do this with two simple functions. Type the following commands in R:  

```{r}
df$Month <- months(DateConvert)
df$Weekday = weekdays(DateConvert)
```

This creates two new variables in our data frame, Month and Weekday, and sets  
them equal to the month and weekday values that we can extract from the Date object.  
Lastly, replace the old Date variable with DateConvert by typing:  
```{r}
df$Date <- DateConvert
```

Using the table command, answer the following questions.

## Q2.3: In which month did the fewest motor vehicle thefts occur?

```{r}
df %>% count(Year, Month, Arrest) %>%
  filter(Arrest == FALSE) %>% 
  arrange(n)
```
Right answer:
```{r}
table(df$Month)
```

## Q2.4: On which weekday did the most motor vehicle thefts occur?
```{r}
table(df$Weekday)
```

Each observation in the dataset represents a motor vehicle theft, and the Arrest  
variable indicates whether an arrest was later made for this theft.  
  
## Q2.5: Which month has the largest number of motor vehicle thefts for which an arrest was made?  

```{r}
aa <- subset(df, df$Arrest == TRUE)
table(aa$Month)
```
The same via tidyverse:
```{r}
df %>% 
  filter(Arrest == T) %>% 
  count(Month, Arrest) %>% 
  arrange(desc(n))
```

## Problem 3.1 - Visualizing Crime Trends  
  
Now, let's make some plots to help us better understand how crime has changed  
over time in Chicago. Throughout this problem, and in general, you can save your  
plot to a file.  
  
First, let's make a histogram of the variable Date. We'll add an extra argument,  
to specify the number of bars we want in our histogram. In your R console, type:
```{r}
hist(df$Date, breaks=100)
```

Looking at the histogram, answer the following questions.  

*Q: In general, does it look like crime increases or decreases from 2002 - 2012?*  
*A:* Decreases  
  
*Q: In general, does it look like crime increases or decreases from 2005 - 2008?*  
*A:* Decrease  
  
*Q: In general, does it look like crime increases or decreases from 2009 - 2011?*  
*A:* Increase  
  
## Problem 3.2 - Visualizing Crime Trends  
  
Now, let's see how arrests have changed over time. Create a boxplot of the variable  
"Date", sorted by the variable "Arrest". In a boxplot, the bold horizontal  
line is the median value of the data, the box shows the range of values between the  
first quartile and third quartile, and the whiskers (the dotted lines extending  
outside the box) show the minimum and maximum values, excluding any outliers (which  
are plotted as circles). Outliers are defined by first computing the difference  
between the first and third quartile values, or the height of the box. This number  
is called the Inter-Quartile Range (IQR). Any point that is greater than the third  
quartile plus the IQR or less than the first quartile minus the IQR is considered  
an outlier.  

```{r}
boxplot(Date ~ Arrest, df)
```

*Q: Does it look like there were more crimes for which arrests were made in the first half of the time period or the second half of the time period? (Note that the time period is from 2001 to 2012, so the middle of the time period is the beginning of 2007.)*
  
*A:* In the second period  
  
## Problem 3.3 - Visualizing Crime Trends  

Let's investigate this further. Use the table function for the next few questions.  
  
*Q: For what proportion of motor vehicle thefts in 2001 was an arrest made?*  
  
Note: in this question and many others in the course, we are asking for an answer  
as a proportion. Therefore, your answer should take a value between 0 and 1.

```{r}
tab33 <- df %>% filter(Year == 2001) %>% {table(.$Arrest)}
tab33[1]
tab33[2]
tab33[2] / sum(tab33)
```

```{r}
df %>%
  filter(Year == 2001) %>%
  {table(.$Arrest)} %>% 
  {c(.[1], .[2], prop = .[2]/sum(.))}
```

## Problem 3.4 - Visualizing Crime Trends  
  
*Q: For what proportion of motor vehicle thefts in 2007 was an arrest made?*

```{r}
df %>%
  filter(Year == 2007) %>%
  {table(.$Arrest)} %>% 
  {tibble(.[1], .[2], prop = .[2]/sum(.))}
```

## Problem 3.5 - Visualizing Crime Trends  
  
*Q: For what proportion of motor vehicle thefts in 2012 was an arrest made?*  

```{r}
df %>%
  filter(Year == 2012) %>%
  {table(.$Arrest)} %>% 
  {tibble(.[1], .[2], prop = .[2]/sum(.))}
```
Since there may still be open investigations for recent crimes, this could explain  
the trend we are seeing in the data. There could also be other factors at play,  
and this trend should be investigated further. However, since we don't know when  
the arrests were actually made, our detective work in this area has reached a dead end.  
  
## Problem 4.1 - Popular Locations  
  
Analyzing this data could be useful to the Chicago Police Department when deciding  
where to allocate resources. If they want to increase the number of arrests that  
are made for motor vehicle thefts, where should they focus their efforts?  
  
We want to find the top five locations where motor vehicle thefts occur. If you  
create a table of the LocationDescription variable, it is unfortunately very hard  
to read since there are 78 different locations in the data set. By using the sort  
function, we can view this same table, but sorted by the number of observations  
in each category. In your R console, type:
```{r}
sort(table(df$LocationDescription), decreasing = T)
```
*Q: Which locations are the top five locations for motor vehicle thefts, excluding the "Other" category? You should select 5 of the following options*  
```{r}
df %>%
  filter(LocationDescription != 'OTHER') %>% 
  count(LocationDescription, sort = T) %>% 
  top_n(5, n)
```

## Problem 4.2 - Popular Locations  
  
Create a subset of your data, only taking observations for which the theft happened  
in one of these five locations, and call this new data set "Top5".  

Create character vector from top 5 LocationDescription's name:
```{r}
(df %>%
  filter(LocationDescription != 'OTHER') %>% 
  count(LocationDescription, sort = T) %>% 
  top_n(5, n) %>% 
  {as.character(.$LocationDescription)} -> Top5_Loc_Desc_Name)
```
Create a subset:
```{r}
(df %>% 
  filter(LocationDescription %in% Top5_Loc_Desc_Name) -> Top5)
```
Check whether subset is correct:
```{r}
Top5 %>% 
  count(LocationDescription) %>% 
  cbind(Top5_Loc_Desc_Name)
```

*Q: How many observations are in Top5?*
```{r}
nrow(Top5)
```

## Problem 4.3 - Popular Locations  
  
R will remember the other categories of the LocationDescription variable from the  
original dataset, so running table(Top5$LocationDescription) will have a lot of  
unnecessary output. To make our tables a bit nicer to read, we can refresh this  
factor variable. In your R console, type:
```{r}
Top5$LocationDescription = factor(Top5$LocationDescription)
levels(Top5$LocationDescription)
```

Use the Top5 data frame to answer the remaining questions.

*Q: One of the locations has a much higher arrest rate than the other locations.*  
*Which is it? Please enter the text in exactly the same way as how it looks in the answer options for Problem 4.1.*
```{r}
Top5_Loc_Desc_Name %>% 
  map_dfr(~ Top5 %>% 
            filter(LocationDescription == .x) %>% 
            {table(.$Arrest)} %>% 
            {tibble(False = .[1], True = .[2], Prop = .[2]/sum(.))}
          )
```
Right answer:
```{r}
Top5 %>%
  filter(LocationDescription == 'GAS STATION') %>% 
  {table(.$Arrest)} %>% 
  {tibble(.[1], .[2], prop = .[2]/sum(.))}
```

## Problem 4.4 - Popular Locations  
  
*Q: On which day of the week do the most motor vehicle thefts at gas stations happen?*  

```{r}
Top5%>%
  filter(LocationDescription == 'GAS STATION') %>% 
  {table(.$Weekday)}
```

## Problem 4.5 - Popular Locations  
  
*Q: On which day of the week do the fewest motor vehicle thefts in residential driveways happen?*

```{r}
Top5 %>%
  filter(LocationDescription == "DRIVEWAY - RESIDENTIAL") %>% 
  count(.$Weekday, sort = T)
```

