---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load library
```{r cars}
WHO <- read.csv('data/WHO.csv')
```

## Look into data
```{r}
str(WHO)
```

## Questions

What is the mean value of the "Over60" variable?

```{r}
mean(WHO$Over60)
```
Which country has the smallest percentage of the population over 60?

```{r}
# which.min(WHO$Over60)
WHO$Country[which.min(WHO$Over60)]
```

Which country has the largest literacy rate?
```{r}
# which.max(WHO$LiteracyRate)
WHO$Country[which.max(WHO$LiteracyRate)]
```

## Data visualization

```{r}
boxplot(WHO$LifeExpectancy ~ WHO$Region, xlab="", ylab="Life Expectancy", main="Life Expectancy of Countries by Region")
```
Use table:
```{r}
table(WHO$Region)
```
## Questions

Use the tapply function to find the average child mortality rate of countries in each region.  
Which region has the lowest average child mortality rate across all countries in that region?
```{r}
tapply(WHO$ChildMortality, WHO$Region, mean)
```

