---
title: "Good Visualizations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(ggmap)
library(reshape2)
```

# VIDEO 3 - Bar Charts

Load our data, which lives in intl.csv
```{r}
intl <- read.csv("intl.csv")
str(intl)
```

We want to make a bar plot with region on the X axis and Percentage on the y-axis.
```{r}
ggplot(intl, aes(x=Region, y=PercentOfIntl)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=PercentOfIntl))
```

First of all, though, what is this stat = "identity"? Well, it's pretty simple. Geometry bar has multiple modes of operation, and stat = "identity" says, use the value of the y variable as is, which is what we want. The height of the bar is the value of the y variable. Now, there are other modes, including one that counts the number of rows for each value of x, and plots that instead. So you can look at the documentation for ggplot to see the different options and how they work. But stat = "identity" is what we want right now.  
  
Make Region an ordered factor. We can do this with the re-order command and transform command. 
```{r}
intl <- transform(intl, Region = reorder(Region, -PercentOfIntl))
str(intl)
```

Make the percentages out of 100 instead of fractions
```{r}
intl$PercentOfIntl <- intl$PercentOfIntl * 100
```

Make the plot
```{r}
ggplot(intl, aes(x=Region, y=PercentOfIntl)) +
  geom_bar(stat="identity", fill="dark blue") +
  geom_text(aes(label=PercentOfIntl), vjust = -0.4) +
  ylab("Percent of International Students") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1))
```

Load in the international student data
```{r}
intlall <- read.csv("intlall.csv",stringsAsFactors=FALSE)
```

Lets look at the first few rows
```{r}
head(intlall)
```

Those NAs are really 0s, and we can replace them easily
```{r}
intlall[is.na(intlall)] <- 0
```

Now lets look again
```{r}
head(intlall) 
```

Load the world map
```{r}
world_map <- map_data("world")
str(world_map)
```

Lets merge intlall into world_map using the merge command
```{r}
world_map <- merge(world_map, intlall, by.x ="region", by.y = "Citizenship")
str(world_map)
```

Plot the map
```{r}
ggplot(world_map, aes(x=long, y=lat, group=group)) +
  geom_polygon(fill="white", color="black") +
  coord_map("mercator")
```

Reorder the data
```{r}
world_map <- world_map[order(world_map$group, world_map$order),]
```

Redo the plot
```{r}
ggplot(world_map, aes(x=long, y=lat, group=group)) +
  geom_polygon(fill="white", color="black") +
  coord_map("mercator")
```

Lets "fix" that in the intlall dataset
```{r}
intlall$Citizenship[intlall$Citizenship=="China (People's Republic Of)"] = "China"
```

We'll repeat our merge and order from before
```{r}
world_map = merge(map_data("world"), intlall, 
                  by.x ="region",
                  by.y = "Citizenship")
world_map = world_map[order(world_map$group, world_map$order),]
```

```{r}
ggplot(world_map, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill=Total), color="black") +
  coord_map("mercator")
```

We can try other projections - this one is visually interesting
```{r}
ggplot(world_map, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill=Total), color="black") +
  coord_map("ortho", orientation=c(20, 30, 0))
```

```{r}
ggplot(world_map, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill=Total), color="black") +
  coord_map("ortho", orientation=c(-37, 175, 0))
```

# VIDEO 7 - Line Charts

Now lets load our dataframe
```{r}
households <- read.csv("households.csv")
str(households)
```

Lets look at the first two columns of our households dataframe
```{r}
households[,1:2]
```

First few rows of our melted households dataframe
```{r}
head(melt(households, id="Year"))
```

```{r}
households[,1:3]
```

```{r}
melt(households, id="Year")[1:10,3]
```

```{r}
melt(households, id="Year")[1:10,]
```

Plot it
```{r}
ggplot(melt(households, id="Year"),       
       aes(x=Year, y=value, color=variable)) +
  geom_line(size=1) + geom_point(size=5) +  
  ylab("Percentage of Households")
```


